---
title: "Estimation with Dummy Independent Variables"
---

```{r, warning= FALSE, message= FALSE}
library(tidyverse)
```

## Sample data

```{r,echo=TRUE}
UCBdata <- UCBAdmissions %>% as_tibble() #Little bit of cleaning
head(UCBdata, 3)
```

Suppose we want to estimate 
$$n = \beta_0 + \beta_1 Admit + \epsilon$$

## Dummy variables in R
There are many ways to estimate with dummy independent variables in R

- Direct estimation with *character vector*
  - Pro: easy to implement (if our data is similar to *UCBAdmissions* data)
  - Con: difficult to change reference level
  
- Create *0-1 variable* for estimation
  - Pro: we can manually specify the reference level 
  - Con: cumbersome if we have many categories

- Coerce the variable to a *factor*
  - Pro: handle variables with many categories well (cf. fixed effects), can be used with numeric vectors
  - Con: can be tricky to handle, change reference level

## Estimation with Charactor Vectors

```{r, echo=TRUE}
lm(data = UCBdata,n ~ 1 + Admit)
```

The coefficient's name displayed in the console follows the format 'variable name' + 'level coded as 1'. Therefore, "AdmitRejected" implies that R has coded Rejected as 1 and Admitted as 0. 

R decides which one is reference level by its alphabetical order.

## Creating 0-1 variables using *ifelse()*

We can also directly create a new dummy variable.

```{r, echo=TRUE}
UCBdata$admitDum <- ifelse(UCBdata$Admit == "Admitted", 1, 0)
lm(data = UCBdata,n~1+admitDum)
```

This process can be cumbersome if we have many categories (cf. fixed effects) or many categorical variables. 

## Creating 0-1 variables using *I()*

```{r, echo=TRUE}
I(UCBdata$Admit=="Admitted")
lm(data = UCBdata,n~1+I(Admit=="Admitted"))
```

## Estimation with Integer and Numerical Vectors

Let's use mtcars data set. Suppose we want to estimate 
$$mpg = \alpha + \beta_1I(cyl=6) +\beta_2I(cyl=8) + \epsilon$$
```{r}
head(mtcars, 3)
```

The following code does not work. Why?
```{r, echo=TRUE}
lm(data = mtcars, mpg~1+cyl)
```

We may use I() function to do the job.

```{r, echo=TRUE}
lm(data = mtcars,mpg~1+I(cyl==6)+I(cyl==8))
```

However, if our variable contains many categories, this may be time-consuming. One easy fix is to coerce the variable into a factor.

## Understanding Factors

A factor looks like a vector, but it has special properties. R keeps track of the unique values in a vector, and each unique value is called a *level* of the associated factor.

If we have a vector of strings or integers, we can create a categorical variable by using the command *factor()*.

```{r, echo=TRUE}
strVec <- c("Win", "Win", "Lose", "Tie", "Win", "Lose")
f <- factor(strVec)
f
```

Notice that when we printed the factor, f, R did not put quotes around the values. They are levels, not strings.

R also displays the distinct levels below the factor.

## Estimation with a Factor Variable

Going back to the previous question, we can coerce cyl into a factor.

```{r, echo=TRUE}
lm(data = mtcars,mpg~1+factor(cyl))
```

This will be useful when we estimate models with many fixed effects.

## Dummy Interaction Variables

```{r, echo=TRUE}
lm(data = mtcars, mpg~1+cyl*hp)
```
```{r, echo=TRUE}
lm(data = mtcars, mpg~1+factor(cyl)*hp)
```

## Reordering Factor

```{r, echo=TRUE}
factor(mtcars$cyl)
```

Originally, 4-cylinder is the base case. Suppose, we prefer 8-cylinder cars to be the base case. We can reorder the factor with *fct_relevel* from the *forcats* package.

```{r, echo=TRUE, warning=FALSE}
library(forcats)
cylf <- mtcars$cyl %>%
  factor() %>%
  fct_relevel(levels = "8") #specify the 1st level
print(cylf)
```

Notice that the data order does not change. Therefore, we don't have to worry about rearranging other variables.

## Estimation with a Releveled Factor

Since the releveled factor was saved in new variable cylf, we need to estimate with the new variable instead.

```{r, echo=TRUE}
lm(mtcars$mpg~1+cylf)
```

Now 8-cylinder car has become the base case.
