---
title: "Missing Values"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

In R, missing values are represented by the symbol NA (not available). Impossible values (e.g., dividing by zero) are represented by the symbol NaN (not a number).

```lm()``` automatically omits missing values. This makes it tricky to know which observations are included in the estimation. To avoid this, we may remove rows that contain any NA values using ```na.omit()```. 

However, ```na.omit()``` may discard more rows than necessary. Suppose we have a data frame that contains $x,y,$ and $z$. We want to regress $y$ on $x$ but $z$ has some missing values. If we use ```na.omit()``` directly on that data frame, we will end up with fewer observations.

It is important to carefully examine data before dropping any observations. Make sure that omitting data makes sense in your context.

In this lesson, we need the following packages.

```{r, echo=TRUE, message = FALSE, warning= FALSE}
library(wooldridge)
library(tidyverse)
```

## Counting NAs

```is.na()``` returns TRUE if that element is missing.

```{r, echo=TRUE}
x <- c(1,2,3,NA,5)
y <- is.na(x)
print(y)
```

Logical values i.e. ```TRUE``` or ```FALSE``` can be treated as numerical values i.e. ```1``` or ```0```. It also has properties of numbers. For example, we can sum over the vector ```y```.

```{r,echo = TRUE}
sum(y)
```

This means we can count the number of missing values in a dataframe using the combination of ```is.na()``` and ```sum()```.

```{r}
discrim$psoda %>%
  is.na() %>%
  sum()
```

This means there are eight missing values in ```discrim$psoda```. Repeat this for all the necessary variables, or to save time, we may put all the columns in a list and loop over it.

```{r, echo=TRUE, message = FALSE, warning= FALSE}
var <- list(discrim$psoda, discrim$prpblck, discrim$income)

for (i in 1:length(var)) {
  var[[i]] %>%
    is.na() %>% 
    sum() %>% #Summing logical values
    print()
}
```

There are eight NAs in $psoda$, one in $prpblck$ and one in $income$.

## Find rows with NAs

We want to extract the index of rows with missing value(s). We may combine ```is.na()``` with ```which()```. Again, loop over the list ```var``` we defined above.

```{r, echo=TRUE}
for (i in 1:length(var)) {
  which(is.na(var[[i]])==TRUE) %>% print()
}
```

```which()``` returns the row number that satisfies the condition in its argument. If we use ```na.omit()``` on a data frame with these three columns, we will discard 9 rows.

## Removing NAs from a Data Frame

Before removing those rows, let's check the number of rows in our data frame.

```{r, echo=TRUE}
rawdata <- discrim %>%
  select(psoda, prpblck, income)

nrow(rawdata)
```

After removing the rows with missing values, confirm the result using ```nrow()``` again.

```{r, echo=TRUE}
cleandata <- na.omit(rawdata)
nrow(cleandata)
```

We have successfully discarded 9 rows.