---
title: "Missing Values"
---

- In R, missing values are represented by the symbol NA (not available). Impossible values (e.g., dividing by zero) are represented by the symbol NaN (not a number).

- *lm()* command automatically omits missing values. (cf. HW4) This is problematic.

- To avoid this, we may remove rows that contain any NA values using *na.omit()*. 

- However, *na.omit()* may discard more rows than necessary. 
  - Suppose we have a data frame that contains $x,y,$ and $z$. 
  - We want to regress $y$ on $x$ but $z$ has some missing values. 
  - If we use na.omit() directly on that data frame, we will end up with fewer observations.

- Carefully examine data before dropping any observations. Make sure that omitting data makes sense in your context.

## Counting NAs

- is.na() returns TRUE if that element is missing.

```{r, echo=TRUE}
x <- c(1,2,3,NA,5)
y <- is.na(x)
print(y)
```

- Logical values i.e. TRUE or FALSE can be treated as numerical values i.e. 1 or 0.

```{r,echo = TRUE}
sum(y)
```

- This means we can count the number of missing values in a dataframe using the combination of *is.na()* and *sum()*.

```{r, echo=TRUE, message = FALSE, warning= FALSE}
library(wooldridge)
library(tidyverse)
var <- list(discrim$psoda, discrim$prpblck, discrim$income)

for (i in 1:length(var)) {
  var[[i]] %>%
    is.na() %>% 
    sum() %>% #Summing logical values
    print()
}
```

- There are eight NAs in $psoda$, one in $prpblck$ and one in $income$.

## Find rows with NAs

- We want to extract the index of rows with missing value(s)

- Combine *is.na()* with *which()*

```{r, echo=TRUE}
for (i in 1:length(var)) {
  which(is.na(var[[i]])==TRUE) %>% print()
}
```

- If we use na.omit() on this data frame, we will discard 9 rows.

## Removing NAs from a Data Frame

- Before removing those rows, let's check the number of rows in our data frame.

```{r, echo=TRUE}
rawdata <- discrim %>%
  select(psoda, prpblck, income)

nrow(rawdata)
```

- After removing the rows with missing values, confirm the result using *nrow()* again.

```{r, echo=TRUE}
cleandata <- na.omit(rawdata)
nrow(cleandata)
```

- We have successfully discarded 9 rows.