---
title: "Probability Distributions in R"
---

Each distribution is indicated by a root name. e.g. root name for normal distribution is *norm*. Other root names are such as *binom*, *chisq*, *f*, *t*, *unif*

Every distribution has four functions
  - *p*: probability (cdf)
  - *q*: quantile (inverse cdf)
  - *d*: density (pdf)
  - *r*: random (random variable having the specified distribution)

## Generating Random Number

Suppose we want to draw 10 values of $\epsilon$ such that $\epsilon$ follows $N(0,1)$.

```{r, echo=TRUE}
rnorm(10, mean = 0, sd = 1) #default is mean = 0, sd = 1
```

## Cumulative distribution function

It is known that 95% of the area under the normal distribution lies within 1.96 standard deviations away from the mean. We can see this by plugging in the cumulative probability to function *q* of root *norm*.

```{r, echo=TRUE}
c <- qnorm(p = c(0.025,0.975))
print(c)
```

We can confirm the cumulative probability of *c* by using function *p*.

```{r, echo=TRUE}
pnorm(c)
```
