---
title: "RESET"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

## Review of RESET

Regression Specification Error Test (RESET) has been used to detect general functional form misspecifications.

Consider 
$$
y = \beta_0 + \beta_1x_1 + ... + \beta_k x_k + u
$$

RESET involves three steps: First, estimate the original model. Obtain the fitted values. Second, estimate a new model with polynomials of the fitted values.
$$
y = \beta_0 + \beta_1x_1 + ... + \beta_k x_k + \delta_1\hat{y}^2 + \delta_2\hat{y}^3+u
$$
Third, test the joint significance of $\delta_1$ and $\delta_2$.

For this exercise, load the following packages.
```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(wooldridge)
library(tidyverse)
library(car)
```

## Example 9.2 Housing Price Equation
Use ```hprice1``` data from the ```Wooldridge``` package. Consider two models for housing prices.
$$
price = \beta_0 + \beta_1 lotsize + \beta_2 sqrft + \beta_2 bdrms + u
$$

$$
lprice = \beta_0 + \beta_1 llotsize + \beta_2 lsqrft + \beta_2 bdrms + u
$$

We want to compare whether the level-level model or the log-log model is preferred.

```{r, echo=TRUE, warning=FALSE, message=FALSE}
hprice <- hprice1

# Step 1
levelmod <- lm(data = hprice,price ~ 1 + lotsize + sqrft + bdrms)
hprice$predsq <- levelmod$fitted.values^2
hprice$predcb <- levelmod$fitted.values^3

# Step 2
levelmod_reset <- lm(data = hprice,
                     price ~ 1 + lotsize + sqrft + bdrms + predsq +predcb)

# Step 3
linearHypothesis(levelmod_reset,c("predsq=0", "predcb=0"))
```

Repeat the same exercise for the log model. For the log model, I write a simple function for RESET as follows. This function takes an ```lmobject``` as input. An advantage of writing a function is that we don't have to copy-paste all the codes. This point will be demonstrated below.

```{r, echo=FALSE}
myresettest <- function(lmobj){
  # Extract data used from the lmobject
  x <- lmobj$model[,-1] %>% as.matrix()
  y <- lmobj$model[,1] %>% as.matrix()

  # Step 1
  predsq <- lmobj$fitted.values^2
  predcb <- lmobj$fitted.values^3

  x <- cbind(x, predsq, predcb) %>% as.matrix()

  # Step 2
  resetmod <- lm(y~x)

  # Step 3
  linearHypothesis(model = resetmod,
                   hypothesis.matrix = c("xpredsq=0", "xpredcb=0"),
                   singular.ok = TRUE)
}
```

Now apply this function to the log model.
```{r}
logmod <- lm(data = hprice,
             lprice ~ 1 + llotsize + lsqrft + bdrms)
myresettest(logmod)
```

We can also apply the function to the level model.
```{r}
myresettest(levelmod)
```

For the level-level model, since p-value is statistically significant at 5% significance level, we reject the null hypothesis that $\delta_1 =0, \delta_2 = 0$. Thus, there is evidence for functional form misspecification.

On the other hand, we do not reject the null hypothesis of log-log model at 5% level (although we would at 10% level). This means the log-log model is preferred on the basis of RESET.

However, a drawback of RESET is that it provides no real direction on how to proceed if the model is rejected. Rejecting the level model does not immediately suggest that the log-log model is the next step.